---
title: "rde Tutorial"
author: "Stefan Kloppenborg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rde Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
csl: ieee-with-url.csl
references:
- id: popclock
  title: Current Population
  author:
  - literal: U.S. Census Bureau
  URL: https://www.census.gov/popclock/print.php?component=counter
  type: webpage
  accessed:
    year: 2018
    month: 3
    day: 13
---

```{r setup, include = FALSE}
library(knitr)
library(rde)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When sharing R Notebooks with others, it's not uncommon for the notebook to reference data that is only available on your machine. It could be that the recipient does not have access to a certain database, or it could be as simple as you forgetting to email them a CSV file with the data. In either of these casese, the analysis in the notebook is not self-contained. The package `rde` solves this problem by allowing you to embed the data direclty in the notebook.

If you're running on an X11 system (ie. Linux, or similar), please read the section on configuring the clipboard below before proceeding.

Let's take an example. Let's say that we have a spreadsheet of populations of the ten most populus countries (data originally taken from [@popclock]). Somewhere near the top of our R Notebook, we have a code chunk that looks like the following (in the code below, we're actually loading data from the `rde` package: you would probably be loading it from some other directory on your computer).

```{r}
fname <- system.file("extdata", "country_pop.csv", package = "rde")
pop.data <- read.csv(fname, stringsAsFactors = FALSE)
```

```{r}
kable(pop.data)
```

Now, if you send yuor notebook to someone else and don't send along the file `country_pop.csv`, that person can look at your notebook, but they won't be able to re-run it. If you want to include the data directly in the notebook, you can use `rde` to do so.

`rde` provides two functions: `load_rde_var` and `copy_rde_var`. You'll use `load_rde_var` in your notebook, and you'll use `copy_rde_var` to create one of the arguments that `load_rde_var` needs.

The function `load_rde_var` takes three arguments. The first argument is a boolean (we'll come back to this). The second argument is `loadFcn`. This is a piece of code that loads data from a source of your choosing (a CSV file, a database, etc.). This is the code that needs to work on your computer: it does not need to work on the computer of the notebook recipient. The third argument is `cache`. This argument is a piece of code that recreates the data. You would normally get the contents of this third argument using the `copy_rde_var` function. We'll come back to that.

When you call `load_rde_var`, the function will first try to load the data using the code in the `loadFcn` argument. If this fails, it will fall back on using the `cache`. In this case, it will give you a message to say that it used the cache instead of loading new data. This is what the recipent of your notebook would see if you neglected to send them the data file.

If `load_rde_var` succeeds in loading the data using the code in `loadFcn`, it will then compare this data with the data in `cache`. If there's a difference, it will give you a warning. If you expected the data to change, you can go ahead and update the third argument (again using `copy_rde_var`); if you didn't expect the data to change, well, now you know that it did.

Now we'll come back to that first argument of `load_rde_var`. This argument is a boolean called `useCache`. This allows you to force `load_rde_var` to load data from the cache instead of running the code in `loadFcn`. Under most circumstances, this should be `FALSE`. However, sometimes, it may take a very long time to load your data from its original source (maybe the code executes a very long running database query, or scrapes a million webpages from the internet and just gives you a summary statistic). In the case that you don't want to wait around while you load the data from its original source again, you can set that first argument to `TRUE` and just use th ecahced data.

Continuing on with our example of loading the populations of the ten most populous countries, we would start by wrapping our existing code inside the second argument of `load_rde_var`. It would now look the this:

```{r}
library(rde)

pop.data <- load_rde_var(
  useCache = FALSE,
  loadFcn = {
    fname <- system.file("extdata", "country_pop.csv", package = "rde")
    read.csv(fname, stringsAsFactors = FALSE)
  },
  cache = {
    # We'll fill this in shortly
  }
)
```

If we run that code as is, it will raise a warning. We would expect this since there is no code in the cache argument, so of course, the result of the `loadFcn` and `cache` are different. We'll need to fill in cache argument of `load_rde_var`.

You'd normally start by loading your data into memory as you normally would (the code above would work fine). Once the data `pop.data` is in memory, you're going to copy it into the `cache` argument of `load_rde_var`. You can use `copy_rde_var` to do so.

In the console, you would type:

```{r eval=FALSE, include=TRUE}
copy_rde_var(pop.data)
```

When you execute this, your clipboard will contain some R code that will recreate the variable. Your clipboard will look like this:

```
structure(list(Country = structure(c(3L, 4L, 10L, 5L, 2L, 8L, 
7L, 1L, 9L, 6L), .Label = c("Bangladesh", "Brazil", "China", 
"India", "Indonesia", "Japan", "Nigeria", "Pakistan", "Russia", 
"United States"), class = "factor"), Population = c(1384688986L, 
1296834042L, 329256465L, 262787403L, 208846892L, 207862518L, 
195300343L, 159453001L, 142122776L, 126168156L)), .Names = c("Country", 
"Population"), class = "data.frame", row.names = c(NA, -10L))
```

You can go ahead and paste that into the `cache` argument of `load_rde_var`. Make sure that you paste it inside a pair of braces. The code at the top of your notebook will now look like this:

```{r}
library(rde)

pop.data <- load_rde_var(
  useCache = FALSE,
  loadFcn = {
    fname <- system.file("extdata", "country_pop.csv", package = "rde")
    read.csv(fname, stringsAsFactors = FALSE)
  },
  cache = {
    structure(list(Country = structure(c(3L, 4L, 10L, 5L, 2L, 8L, 
      7L, 1L, 9L, 6L), .Label = c("Bangladesh", "Brazil", "China", 
      "India", "Indonesia", "Japan", "Nigeria", "Pakistan", "Russia", 
      "United States"), class = "factor"), Population = c(1384688986L, 
      1296834042L, 329256465L, 262787403L, 208846892L, 207862518L, 
      195300343L, 159453001L, 142122776L, 126168156L)), .Names = c("Country", 
      "Population"), class = "data.frame", row.names = c(NA, -10L))
  }
)
```

Now, when we run this, it won't raise a warning because `loadFcn` and `cache` are the same.

If you send this notebook to someone else, but neglect to send the data file, they can now still play around with the data because it's now directly in the code. They will, however, get a message indicating that the data has been loaded from cache.

What if you inadvertently change the data file? Or if you're reading the data from a database that changes? Well, if that happens, `loadFcn` and `cache` won't match. In this case, you'll get a warning. This can be useful: maybe you didn't expect the data to change, or maybe you need to update some of the text in your notebook --- maybe some of your conclusions or explanation needs to change. Assuming that the change in the data file (or database) isn't some sort of mistake, make sure that you update the value of the `cache` argument with the new data (again, you'll use the `copy_rde_var` function to do so).

# Installing on X11 Systems
If you're on an X11 system (like Linux), you'll need to install some additional software. You should not have to do this on Windows or Mac. On X11 systems, you'll need to install either `xsel` or `xclip`. Depending on the distribution that you use, you will probably install it using a command like `sudo apt-get install xsel`


# References

